Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-05-27 manuscript/manuscript.Rmd
ms updates

We used random draws from the final 1,000 iterations from each of three MCMC chains to introduce stochasticity into our population models. At each time step, we randomly selected climate covariates from one of the 14 observed years. Then, we drew the full parameter set (climate effects and density-dependence fixed effects) from a randomly selected MCMC iteration. Since our focus was on the contribution of climate covariates to population states, we set the random year effects and the random group effects to zero.

@@

Results
-------
### Comparison of Forecast Models

### Forecasting Climate Change Impacts

@@

1. Climate detection
2. Limitations:
  * Sample size at quadrat level
  * are these species inherently sensitive to climate (compare to Chu and Adler in review -- say why we chose Montana site -- most sensitive to climate variables)
  * we know recruitment is important, and this is something that may be easily missed at quad level
3. Implications

>>>>

We used random draws from the final 1,000 iterations from each of three MCMC chains to introduce stochasticity into our population models. At each time step, we randomly selected climate covariates from one of the 14 observed years. Then, we drew the full parameter set (climate effects and density-dependence fixed effects) from a randomly selected MCMC iteration. Using this approach, rather than simply using coefficient point estimates, ensures that relatively unimportant climate covariates (those that broadly overlap 0) have little effect on the simulation results. Since our focus was on the contribution of climate covariates to population states, we set the random year effects and the random group effects to zero.

@@

Results
-------
### Comparison of Forecast Models
We performed one-step-ahead forecasts to compare the two models (IPM vs. QBM). We did not use the random year effects for these forecasts. Thus, beyond the effect of genet size or quadrat cover, only the climate covariates could affect expect plant cover. The models had similar accuracy ($\rho$, correlation between observations and mean prediction from 100 forecasts) across all species (Table 1). However, the IPM had significantly lower overall error (MAE, mean absolute error) for two species (*H. comita* and *P. smithii*), and in no case did the QBM significantly outperform the IPM (Table 1).

### Forecasting Climate Change Impacts
Equilibrium cover simulated from the models was sensitive to climate perturbations, but the IPM and QBM produced inconsistent results (Figure 3). For *B. gracilis*, the IPM predicted a modest increase in cover with a 1% increase in the mean of precipitation or a 1% increase in the mean of temperature, the compounding effect of both being a 20% increase in cover. This reflects the relatively strong effects of precipitation and climate on *B. gracilis* genet growth and recruitment (Figure 1). The QBM also predicted increased *B. gracilis* cover with a precipitation increase, but increasing temperature decreased equilibrium cover (Figure 3).

The IPM and QBM produced consistent predictions for *H. comita* under increased precipitation and when both precipitation and temperature were increased (Figure 3). However, the IPM predicted more modest changes than the QBM, and for a temperature increase the two models differed: the IPM predicted an increase in cover while the QBM predicted the opposite.

@@

It is perhaps startling that two different types of models fit using the same data produce drastically different forecasts (Figure 3).

### 2015-05-27 manuscript/manuscript.tex
ms updates

@@

@@

@@

\begin{table}[ht]
\centering
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & Accuracy & Mean Obs. Cover \\ 
  \hline
IPM & BOGR & 4.77 & 0.90 & 9.16 \\ 
  QBM & BOGR & 5.30 & 0.88 & 9.22 \\ 
  IPM & HECO & 0.60 & 0.80 & 1.22 \\ 
  QBM & HECO & 2.03 & 0.84 & 1.25 \\ 
  IPM & PASM & 0.18 & 0.65 & 0.40 \\ 
  QBM & PASM & 0.17 & 0.87 & 0.40 \\ 
  IPM & POSE & 0.78 & 0.61 & 1.23 \\ 
  QBM & POSE & 1.30 & 0.57 & 1.25 \\ 
   \hline
\end{tabular}
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation.} 
\end{table}

>>>>

We performed one-step-ahead forecasts to compare the two models (IPM
vs.~QBM). We did not use the random year effects for these forecasts.
Thus, beyond the effect of genet size or quadrat cover, only the climate
covariates could affect expect plant cover. The models had similar
accuracy ($\rho$, correlation between observations and mean prediction
from 100 forecasts) across all species (Table 1). However, the IPM had
significantly lower overall error (MAE, mean absolute error) for two
species (\emph{H. comita} and \emph{P. smithii}), and in no case did the
QBM significantly outperform the IPM (Table 1).

@@

Equilibrium cover simulated from the models was sensitive to climate
perturbations, but the IPM and QBM produced inconsistent results (Figure
3). For \emph{B. gracilis}, the IPM predicted a modest increase in cover
with a 1\% increase in the mean of precipitation or a 1\% increase in
the mean of temperature, the compounding effect of both being a 20\%
increase in cover. This reflects the relatively strong effects of
precipitation and climate on \emph{B. gracilis} genet growth and
recruitment (Figure 1). The QBM also predicted increased \emph{B.
gracilis} cover with a precipitation increase, but increasing
temperature decreased equilibrium cover (Figure 3).

The IPM and QBM produced consistent predictions for \emph{H. comita}
under increased precipitation and when both precipitation and
temperature were increased (Figure 3). However, the IPM predicted more
modest changes than the QBM, and for a temperature increase the two
models differed: the IPM predicted an increase in cover while the QBM
predicted the opposite.

@@

It is perhaps startling that two different types of models fit using the
same data produce drastically different forecasts (Figure 3).

@@

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & Accuracy & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 4.77 & 0.90 & 9.16 \\ 
  BOGR & QBM & 5.30 & 0.88 & 9.22 \\ 
  HECO & IPM & 0.60 & 0.80 & 1.22 \\ 
  HECO & QBM & 2.03 & 0.84 & 1.25 \\ 
  PASM & IPM & 0.18 & 0.65 & 0.40 \\ 
  PASM & QBM & 0.17 & 0.87 & 0.40 \\ 
  POSE & IPM & 0.78 & 0.61 & 1.23 \\ 
  POSE & QBM & 1.30 & 0.57 & 1.25 \\ 
   \hline
\end{tabular}
\end{table}

The IPM MAE is significantly lower for HECO ($p = 3.3 \times 10^{-10}$)
and POSE ($p = 0.00013$). MAEs are statisticially similar between models
for BOGR and PASM.
